Analyzing the impact of social attributes on the success of commit integration

### 2017-02-19 Paper/MSRChallenge2017.tex
addressed the other comments

We present the approach taken to guide this study by
following the Goal-Question-Metric (GQM) methodology descibed by Basili et 
al~\cite{Basili84}.

@@

\begin{comment}
we have created a dataset with technical attributes to be able to analyze the predictiveness of these attributes and its relationship with commit integration success. This way, we are able to create a model
We are afterwards interested in knowing if adding social attributes to this 
predictability model will increase the probability of the model asserting 
correctly future instances by knowing not only technical data about the commit 
being pushed but also social attributes of the developer pushing the changes.
\end{comment}

\paragraph{Metric}
\clg{I think I need a little bit more detail here, at multiple levels (like, why
  a decision tree? among other questions.}
We created a decision tree based on two different sets of data. Merely 
technical data related to technical aspects of the complexity of the commit, and 
a superset of this before-mentioned data and social data gathered from the 
developer who pushed the commit. We evaluate the predictive power of these two 
models with 10 fold cross validation and we analyze which of them is able to 
predict with higher accuracy the success of unseen commits.

@@

We also added one item from the original data set provided by the MSR Challenge to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

@@

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check? RESPONSE: Spot check that they did belong to those years? Yeah}

@@

To avoid this and have a 1:1 ratio successful to not successful, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).\clg{How?}  We used both the
downsampled and non-downsampled data sets in our results to show the difference of the behavior of the decision tree when it is trained with a equal ratio than when it is trained with a non equal ratio.

@@

\hline
\end{tabular}
\\
\center
  \caption{Comparison of predictive power by running 10
    fold cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}}   
\end{table*}

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using ten fold cross 
validation with the data in each data set.
To give an answer to main research question: ``Do social attributes, gathered 
from the Github profiles of developers,
allow for more accurate predictions than only using technical data?", we present the following results:

@@

On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, when predicting a successful build in
all instances would produce 50.00\% accuracy.  After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data

@@

For the build success investigation,\clg{Is this a different investigation than
  the previous subsection?}  we classified the build success rate for
core team members and non-core team members.  We found that core team members
triggered of 89.39\% of the builds in the data set and 75.58\% of those builds
passed.  Non-core team members triggered 10.61\% of the builds in the\clg{which?} data set
and 64.70\% of the builds passed.

\clg{upshot?}

@@

We also investigated if developers with followers were more likely to have a
successful build integration, since have followers may be seen as vouching for
the developer's abilities.  We specifically investigated if a developer with 
followers was more likely to pass the build.  We found that only 0.87\% of the
contributions in the data set was performed by developers with zero 
followers (excluding the developers in which we were not able to extract profile
information for). We found that 99.13\% of the contributions were performed by
developers with at least one or more followers.
We found that developers with 0 followers had a successful build commit 68.49\%
of the time.  Developers with one or more followers were successful 74.78\% of
the time.

We also investigated developers with a large number of followers (1000 or more
followers).  We found that 6.94\% of the contributions in the data set were
performed by developers with 1000 or more followers.  We found that these
contributions had a 78.61\% success rate.

\clg{upshot?}

@@

\clg{connection back to theory?}

@@

\paragraph{Construct validity}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@

Our results show that adding social data to the construction of the decision tree increases the probability this correctly predicting future commit's success in integrating into their corresponding projects by a margin of 4.09\% to 11.99\%, depending on the ratio of successful vs unsuccessful instances are used to train the desicion tree.

>>>>

We present the approach taken to guide this study by
following the Goal-Question-Metric (GQM) methodology described by Basili et 
al.~\cite{Basili84}.

@@

\begin{comment}
we have created a dataset with technical attributes to be able to analyze the 
predictive power of these attributes and its relationship with commit integration 
success. This way, we are able to create a model
We are afterwards interested in knowing if adding social attributes to this 
predictability model will increase the probability of the model asserting 
correctly future instances by knowing not only technical data about the commit 
being pushed but also social attributes of the developer pushing the changes.
\end{comment}

\paragraph{Metric}
Because decision trees are well tested and easy to interpret, we used a decision
tree classifier to create models which predict build integration success. 
The first model was generated with only
the technical factors, that provide insight on the complexity of the commit, 
to understand how well technical factors predicted build
success.  We then added the social factors to the technical factors to determine
the influence of additional social factors on predicting build success.
We evaluate the predictive power of these two 
models with 10-fold cross validation and determine the effects of adding social
factors when predicting build integration success.

@@

We also added one item from the MSR Challenge data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

@@

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check? Mau's Response: Spot
check that they did belong to those years? Yeah. Zack's Response: assuming you
are asking a different question of if the values changed in the last three
years, than no, we did not check because we weren't sure how.  What should we
say here?}

@@

To avoid this and have a 1:1 ratio successful to not successful, we created a 
separate data set where the
successful builds were a random sub-sample of successful builds were selected
such that the number of successful builds equaled the number of failing builds,
known as downsampling.  We used 
both the downsampled and non-downsampled data sets in our results to show the 
difference of the behavior of the decision tree when it is trained with a equal 
ratio than when it is trained with a non equal ratio.

@@

\hline
\end{tabular}
\\
\center
  \caption{Comparison of predictive power by running 10-fold
    cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}}   
\end{table*}

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using 10-fold cross 
validation with the data in each data set.
To give an answer to main research question: ``Do social attributes, gathered 
from the Github profiles of developers,
allow for more accurate predictions than only using technical data?", we 
present the following results:

@@

On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, when predicting a successful build in
all instances would produce 50.00\% accuracy.  After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
social data can be used to improve predictions of build integration success.

@@

To determine if core team member's have greater build success rates,
we classified the build success rate for
core team members and non-core team members in the data set without 
downsampling.  We found that core team members
triggered of 89.39\% of the builds in the data set and 75.58\% of those builds
passed.  Non-core team members triggered 10.61\% of the builds in the data set
and 64.70\% of the builds passed.  This shows that core team members are more
likely to have a successful build integration than non-core team members.

@@

We also investigated if developers with followers were more likely to have a
successful build integration, since have followers may be seen as vouching for
the developer's abilities.  We specifically investigated if a developer with 
followers was more likely to pass the build in the data set without 
downsampling.  We found that only 0.87\% of the
contributions in the data set was performed by developers with zero 
followers (excluding the developers in which we were not able to extract profile
information for). We found that 99.13\% of the contributions were performed by
developers with at least one or more followers.
We found that developers with 0 followers had a successful build commit 68.49\%
of the time.  Developers with one or more followers were successful 74.78\% of
the time.

We also investigated developers with a large number of followers (1000 or more
followers).  We found that 6.94\% of the contributions in the data set were
performed by developers with 1000 or more followers.  We found that these
contributions had a 78.61\% success rate.  These results show that developers
with more followers are more likely to have a successful integration build.

@@

Overall, this shows that while experience is not a perfect indicator, more
experienced developers are more likely to make high quality project commits.

@@

\paragraph{Construct validity}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 committed does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing Travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@

Our results show that adding social data to the construction of the decision 
tree increases the probability this correctly predicting future commit's success 
in integrating into their corresponding projects by a margin of 4.09\% to 
11.99\%, depending on the ratio of successful vs unsuccessful instances are 
used to train the decision tree.
